---
title: "RExercise#2_Fantilanan.Rmd"
author: "Glen Mark F. Fantilanan"
date: "2026-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# TODO: Store 17 in a variable called a, and 5 in a variable called b
# Then print the results of: a + b, a - b, a * b, a / b
a <- 17
b <- 5

print(a + b)
print(a - b)
print(a * b)
print(a / b)

# TODO: Using the same a and b, compute and print a %% b and a %/% b
# a %% b = remainder of a / b
# a %/% b = whole-number (integer) division
print(a %% b)
print(a %/% b)

# TODO: Print 2 raised to the power of 10 using the ^ operator
print(2 ^ 10)
```
(a / b) exactly divides them which gives 3.4, (a %% b) which gives the remaining number of 17 which is 2 and (a %/% b) gives exactly the same number of 3 
Yes, both predictions were correct 10 %% 3 = 1 and 10 %/% 3 = 3
```{r}
x <- 42
y <- 42L
z <- "42"
w <- TRUE
v <- 3 + 21

# TODO: Use typeof() to check the type of each of x, y, z, w, v
cat("\n=== TYPE ===\n")
print(typeof(x))
print(typeof(y))
print(typeof(z))
print(typeof(w))
print(typeof(v))

# TODO: Use class() to check the class of each of x, y, z, w, v
cat("\n=== CLASS ===\n")
print(class(x))
print(class(y))
print(class(z))
print(class(w))
print(class(v))

# TODO: Use mode() to check the class of each of x, y, z, w, v
cat("\n=== MODE ===\n")
print(mode(x))
print(mode(y))
print(mode(z))
print(mode(w))
print(mode(v))

```
x and y both contain the number 42, the 42 is treated as a double, while the L in 42L tells R that it should be an integer because it has a special suffix for letter "L"

Yes, they can be different. For example, x has a typeof() of "double", but its class() and mode() are "numeric". This is because the three functions describe the object in different ways. For a matrix or factor, I would also expect them to give different results because a matrix or factor has its own class.
```{r}
num_string <- "123"

num <- as.numeric(num_string)

# TODO: Convert num_string to a numeric value using as.numeric()
# Then check its type with typeof(), class(), and mode()
cat("\n=== NUMERIC ===\n")
print(num)
print(typeof(num))
print(class(num))
print(mode(num))

# TODO: Convert the number 0 and the number 1 to logical using as.logical()
# What do you get?
cat("\n=== TRUE OR FALSE -LOGICAL- ===\n")
print(as.logical(0))
print(as.logical(1))

# TODO: Try converting bad_string to numeric using as.numeric()
# Run this chunk and observe the warning message and the result
cat("\n=== CONVERT ===\n")
bad_string <- "hello123"

print(as.numeric(bad_string))



```
I think R gives a warning because the conversion is not completely impossible for the program to continue. It gives NA for the value that cannot be converted and warns us that something went wrong.


The problem is that I might not notice that a value became NA. This could cause wrong results in later calculations without me realizing that something went wrong.


as.logical(2)
# TRUE

as.logical(-1)
# TRUE

R seems to treat 0 as FALSE and any non-zero number as TRUE. That is why both 2 and -1 become TRUE.
```{r}
first_name <- "Juan"
last_name <- "Dela Cruz"

# TODO: Print first_name and last_name using print() and/or cat()
cat("\n=== 1 ===\n")
print(first_name)
print(last_name)

# TODO: Use paste() to combine first_name and last_name into a full name
# TODO: Use paste0() to do the same, and compare the two results
cat("\n=== 2 ===\n")
full_name <- paste(first_name, last_name)
print(full_name)

full_name2 <- paste0(first_name, last_name)
print(full_name2)

# TODO: Use paste() with the argument sep = "-" to join first_name and
# last_name with a dash instead of a space
cat("\n=== 3 ===\n")
full_name3 <- paste(first_name, last_name, sep = "")
print(full_name3)

cat("\n=== 4 ===\n")
age <- 21
# TODO: Use paste0() to build the sentence:
# "Juan Dela Cruz is 21 years old." by combining full_name, age, and text

sentence <- paste0(first_name, " ", last_name, " is ", age, " years old.")
print(sentence)

```
paste() and paste0() are different because paste() uses a space between the strings by default. paste0() does not use a separator, so it puts the strings directly beside each other.

No, it did not cause an error. paste0() can combine numbers and text together. It automatically converts the number age into a character when joining it with the other strings. This is similar to the type conversion we practiced in Part C because the number is changed into text when it is combined with the character strings.
```{r}
# TODO: Now fix the code above so that "5" + 5 correctly evaluates to 10
# using an appropriate conversion function.
cat("\n=== fix ===\n")
print(as.numeric("5") + 5)

# TODO: Predict, then test: what is the result of "5" + 5 in R?
# Run it and read the error message carefully.
#print("5" + 5)#

```
I think R requires an explicit conversion because it does not want to guess what the programmer means. A value like "5" is a character, even though it looks like a number. Converting it first makes it clear that we want R to treat it as a number.

The numbers were probably read from the text file as characters instead of numeric values. I would first check the data using typeof(), class(), or mode() to see what type it is. If it is character, I would convert it using something like as.numeric(). I would also check for warnings or values that are not actually numbers, because those could cause problems when using sum().
I learned that R treats different types of data differently, such as numbers, characters, logical values, and integers. I also learned that typeof(), class(), and mode() can help identify what type of data I am working with. Understanding data types is important because I may need to convert the data before doing calculations or combining different values. This can help prevent errors in my code.
